Gray-Box Optimization using the Parameter-less Population Pyramid

### 2015-01-21 Gray Box Optimization using the Parameter-less Population Pyramid/main.tex
Abstract to 197 words

Unlike black-box optimization problems, gray-box optimization problems have known,
limited-cardinality relationships between variables. Such problems are
well-known and exist in many real world problems (network security, computational biology, VLSI design, and statistical physics). These
problems are amenable to local search using a Hamming-Ball Hill Climber
(HBHC) that can efficiently find r-bit local optimal. We show how a
simple memetic algorithm in conjunction with HBHC can find \emph{global}
optima for gray-box problems. Finally a gray-box version of P3, which
utilizes the information about variable relationships given by the gray-box
problem itself, is the most efficient of the examined algorithms.

We provide experimental evidence on both NKq-Landscapes and Ising Spin Glasses
that Gray-Box P3 is effective at finding the global optima even for problems with
thousands of variables. This capability is complemented by its efficiency, with
running time and memory usage decreased by a linear factor from Black-Box P3.
On NKq this results in a x375 speedup for problems with at least 1000 variables.
While local search introduces a parameter, we experimentally show this parameter
can likely be fixed to 1 without significant loss in search quality.

\end{abstract}

>>>>

Unlike black-box optimization problems, gray-box optimization problems have known,
limited-cardinality non-linear relationships between variables. This domain contains many
real world problems with examples in network security, computational biology, VLSI design, and statistical physics.
Leveraging these restrictions, the Hamming-Ball Hill Climber
(HBHC) can efficiently find r-bit local optimal. We show how a
simple memetic algorithm in conjunction with HBHC can find \emph{global}
optima for gray-box problems. We also develop a gray-box version of the Parameter-less
Population Pyramid (P3) utilizing both the HBHC and the known information about variable relationships
which outperforms all of the examined algorithms

We provide experimental evidence on both NKq-Landscapes and Ising Spin Glasses
that Gray-Box P3 is effective at finding the global optima even for problems with
thousands of variables. This capability is complemented by its efficiency, with
running time and memory usage decreased by up to a linear factor from Black-Box P3.
On NKq this results in a x375 speedup for problems with at least 1000 variables.
While local search introduces a parameter, we experimentally show this parameter
can likely be fixed to 1 without significant loss in search quality.
\end{abstract}
